---
title: "RDPA1"
author: "iris"
date: "9/20/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##Loading and preprocessing the data
Simply loaded the data into R.
```{r loading data}
data<-read.csv("~/DSC2020/RDPA1/activity/activity.csv")
```

##What is mean total number of steps taken per day?

I calculate the sum of the steps taken each day, plot, then calculate the mean and median of each day. 

```{r steps per day, cache=TRUE, fig.height=3, fig.width=3}
c<-gl(61,288)
sums<-tapply(data$steps, c,sum, na.rm =TRUE)
hist(sums, col = "lightblue", main = "Steps Taken Daily", xlab= "Steps", ylab = "Frequency")
mean(sums)
median(sums)
```

As you can see, the mean is just under 10,000 steps per day.

##What is the average daily activity pattern?

I calculated the average activity level at each interval, then graphed.
```{r steps per interval, cache=TRUE, fig.height=3, fig.width=4}
b<-gl(288,61)
avgint<-tapply(data$steps, b,sum, na.rm =TRUE)
plot(data[1:288,3], avgint,type = "l", col = "limegreen", main = "Average Daily Activity", ylab = "Average Steps", xlab = "Interval")
max(avgint)
```
The highest activity level is at 2100 seconds into the day. 

##Inputing missing values
```{r missing values, fig.height=3,fig.width=4}
sum(is.na(data$steps))
eightdays<-tapply(data$steps, c,sum)
#I counted which days had missing values.
missing<-c(1,8,32,35,40,41,45,61)
B<-missing*288
A<-B-287
#These lists allowed me to see which values needed to be replaced and their location.
list<-c(1:288,2017:2304,8929:9216,9793:10080,11233:11520,11521:11808,12673:12960,17281:17568)
newdata<-replace(data$steps, list, avgint)
newdata<-data.frame(cbind(newdata,data$date,data$interval))
newsums <-tapply(as.integer(newdata$newdata), c,sum)
hist(newsums, col = "pink", main = "Steps Taken Daily (Missing Values Added)", xlab= "Steps", ylab = "Frequency")
mean(newsums)
median(newsums)
```
Adding in the missing data brought the average up.


##Are there differences in activity patterns between weekdays and weekends?

``` {r weekdays}
weekdays<-weekdays(as.Date(newdata$V2))
newdata<-cbind(newdata,weekdays)
mon<-newdata$weekdays == "Monday"
tues<-newdata$weekdays == "Tuesday"
wed<-newdata$weekdays == "Wednesday"
thurs<-newdata$weekdays == "Thursday"
fri<-newdata$weekdays == "Friday"
sat<-newdata$weekdays == "Saturday"
sun<-newdata$weekdays == "Sunday"
weekday<-rbind(newdata[mon,],newdata[tues,],newdata[wed,],newdata[thurs,],newdata[fri,])
weekend<-rbind(newdata[sat,],newdata[sun,])
g<-gl(288,45)
h<-gl(288,16)
avgintweekday<-tapply((as.integer(weekday[,1])), g , sum, na.rm =TRUE)
avgintweekend<-tapply((as.integer(weekend[,1])), h, sum, na.rm =TRUE)
par(mfrow =c(2,1))
plot(weekday[1:288,3], avgintweekday,type = "l", col = "skyblue", main = "Average Weekday Activity", ylab = "Average Steps", xlab = "Interval")
plot(weekend[1:288,3], avgintweekend,type = "l", col = "skyblue", main = "Average Weekend Activity", ylab = "Average Steps", xlab = "Interval")
```
The times that people are active are very different between the weekdays and weekends.

